---
title: "Hemorrhage Detection"
author: "Lilian De Rivera / Swetha Kalla"
date: "4/27/2020"
output: 
 ioslides_presentation:
    css: project.css
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(dplyr)
library(readxl)
library(stringr)
library(tidyr)
library(shiny)
library(ggplot2)
library(DT)
library(reshape)
library(caret)
library(knitr)
library(oro.dicom)
library(pROC)
library(rsconnect)
```
```{r}
# loading the datasets
cu <- read_excel("labels_V2.xlsx")
# summary
any <- cu %>% filter(any==1)
notany <- cu %>% filter(any == 0)
epidural <- cu %>% filter(epidural==1 )
intraparenchymal <- cu %>% filter(intraparenchymal==1 )
intraventricular <- cu %>% filter(intraventricular==1 )
subarachnoid <- cu%>% filter(subarachnoid==1 )
subdural <- cu%>% filter(subdural==1 )
# 
annot1 = cu %>% filter(epidural + intraparenchymal+ intraventricular + subarachnoid + subarachnoid == 1 )
annot2 = cu %>% filter(epidural + intraparenchymal+ intraventricular + subarachnoid + subarachnoid > 1 )
annot0 = notany
```
```{r}
res <- c( nrow(any), nrow(epidural), nrow(intraparenchymal), nrow(intraventricular), nrow(subarachnoid), nrow(subdural) )
tit_res <- c('None','Epidural', 'Intraparenchymal', 'Intraventricular', 'Subarachnoid', 'Subdural')
annot_x <- c(nrow(annot0), nrow(annot1), nrow(annot2) )
tit_annot <- c('0 Annot.','1 Annot.','+1 Annot.')

summarytable <- matrix(c(nrow(cu), nrow(epidural), nrow(intraparenchymal), nrow(intraventricular), nrow(subarachnoid), nrow(subdural), nrow(annot0), nrow(annot1), nrow(annot2)), ncol=1, byrow =TRUE)

rownames(summarytable)<-c('Total Rows', 'Epidural', 'Intraparenchymal', 'Intraventricular', 'Subarachnoid', 'Subdural', '0 Annotation', '1 Annotation', '+1 Annotation')

colnames(summarytable)<-c('No.Images')

summarytable <- as.table(summarytable)

```
```{r}

# Resutls for Model 1

tloss <- read.csv('train_loss_hist1.csv')
vloss <- read.csv('valid_loss_hist1.csv')
tbloss <- read.csv('train_loss_hist_item1.csv')
vbloss <- read.csv('valid_loss_hist_item1.csv')
target <- read.csv('target_labels1.csv')
predicted <-read.csv('pred_labels_per1.csv')
```
```{r message=FALSE, warning=FALSE, paged.print=FALSE}

# Results for Model 2

tloss2 <- read.csv('train_loss_hist2.csv')
vloss2 <- read.csv('valid_loss_hist2.csv')
tbloss2 <- read.csv('train_loss_hist_item2.csv')
vbloss2 <- read.csv('valid_loss_hist_item2.csv')
target2 <- read.csv('target_labels2.csv')
predicted2 <-read.csv('pred_labels_per2.csv')

```


## Introduction
### Machine Learning in Medicine

- Tele - Medicine
- Image Recognition
    - Diagnostics on Cancer Cells
    - Diagnostics on MRI Images
        - **Pulmonary deseases**
        - **Cerebral Tumors** 
    

## Hemorrhage Detection and Subtypes
```{r}
renderImage({
   list(src = "Hemorrhage.png", contentType = 'image/png',width = 750, height = 500,
               alt = "This is alternate text")
    }, deleteFile = FALSE)
  

```


## Problem Definition

**Classify the type of tumor in a MRI Image base on the images presented in the trainning set**

## Project Planning 

```{r}
renderImage({
   list(src = "Project Planning.png", contentType = 'image/png',width = 800, height = 500,
               alt = "This is alternate text")
    }, deleteFile = FALSE)
  

```


## EDA IMAGES

```{r}
ui <- fluidPage(
  tabsetPanel(
    tabPanel("Annotated Images",
             plotOutput("plot1")
             ),
    tabPanel("Images by Type",
             plotOutput("plot2")
             ),
    tabPanel("Summary Images",
             tableOutput("table1")
             ),
    tabPanel("Brain Tumors",
                imageOutput("brain")
        )
  )
)
server <- function(input, output, session){
    output$plot1 <- renderPlot({
         b<-barplot(annot_x/1000, names.arg = tit_annot, cex.names=0.8, las=2)
         text(x=b, y= 150, labels=as.character(annot_x))
    })
    output$plot2 <- renderPlot({
         b<-barplot(res/1000, names.arg = tit_res, cex.names=0.8, las=2)
         text(x=b, y= 5, labels=as.character(res))
    })
    output$table1<- renderTable ({
     
       summarytable
    })
    
    output$brain <- renderImage({
           list(src = "Brainfinal.png", contentType = 'image/png',width = 700, height = 500,
               alt = "This is alternate text")
    }, deleteFile = FALSE)

  

    
    }
shinyApp(ui = ui, server = server)


```

## Research

- **A CNN trainned on a finer disease classes perform better than done trained ddirectly on several classess.**

*Dermatologist- level classification of skin cancer with deep neural networks. by A. Esteva, B. Kuprel, and others.(2017) doi:10.1039/nature2106. page 116.*


- **Networks that include Spatial Transformation Networks can transform regions to a canonical, expected pose to simplify recognition in the following layer.**

*Spatial Trannsformed Networks by M.Jadergerg, K. Simonyan, and others. Google DeepMind UK.(2016)*

##  Prototype Data Augmentation  ( Scope )

- Only a subset of images will be used for the prototype
    - 500 Images per class
    - 1000 Images annotated as with tumor 
    - balanced dataset for training
    - 75% training 25% por validation
    
- A subset of balance dataset will be used for testing 
    - 100 images per class
    - 200 images annotated with tumor and no tumor
    
## Model Architecture
```{r}
ui <- fluidPage(
  tabsetPanel(
    tabPanel("Ensemble_1",
             imageOutput("image1")
             ),
    tabPanel("Ensemble_2",
             imageOutput("image2")
             ),
    tabPanel("STN",
             imageOutput("STN")
             ),
    tabPanel("STN-Transformation",
             imageOutput("STNR")
             )
  )
)
server <- function(input, output, session){
    output$image1 <- renderImage({
    
          list(src = "model1ml2v2.png", contentType = 'image/png',width = 750, height = 450,
               alt = "This is alternate text")
    }, deleteFile = FALSE)

    output$image2 <- renderImage({
    
          list(src = "model2ml2.png", contentType = 'image/png',width = 750, height = 450,
               alt = "This is alternate text")
    }, deleteFile = FALSE)
    
    output$STN <- renderImage({
    
          list(src = "STN.png", contentType = 'image/png',width = 700, height = 425,
               alt = "This is alternate text")
    }, deleteFile = FALSE)

    output$STNR <- renderImage({
    
          list(src = "TransformedImages.png", contentType = 'image/png',width = 750, height = 450,
               alt = "This is alternate text")
    }, deleteFile = FALSE)
    
    }
shinyApp(ui = ui, server = server)

```

## Prototype

- **pyTorch** is used the models and deliver the metrics
    - Convenience due to recent use 
- **AWS** is used as a GPU and python providers
    - Avalilability
- **Data Loader** that reads into memory a batch of images
- **Shiny** is used to process the metrics and visualizations
    - Easy to feed and faster to generate results
- **Pycharm** is used as a python editor
- **vGG16** model for transfer learning, only 16 layers
    - If the model works as expected then it will be changed to *inceptionv3* for phase II


## Model-1 Metrics ( Single Label) 

```{r}

ui <- fluidPage(
  tabsetPanel(
    tabPanel("Loss Function",
             splitLayout(
               plotOutput("plot1")
               ,
               plotOutput("plot2")          
               )
             
             ),
    tabPanel("Loss by Epoch",
             sidebarLayout(
              sidebarPanel(
                selectInput("epoch","No.Epoch", c(1,2,3,4,5,6,7,8,9,10,11,12,13,14),multiple = FALSE)
              ),
              mainPanel(
                splitLayout(
                  plotOutput("plot3"),
                  plotOutput("plot4")
                   )
                )
             )
             
        ),
     tabPanel("Accuracy",
         sidebarLayout(
           sidebarPanel(
             selectInput("epoch2","No.Epoch", c(1,2,3,4,5,6,7,8,9,10,11,12,13,14),multiple = FALSE)
           ),
           mainPanel(
             
              splitLayout(
                  plotOutput("plot5"),
                  plotOutput("plot6")
                   )
           )
         )
               
      ),
    tabPanel("Test Results",
        verticalLayout(
          fluidRow(
            
            column(width=4,
                   plotOutput("roc1")
                   ),
            column(width=4,  
                   plotOutput("roc2")
                   ),
            column(width=4,
                   plotOutput("roc3")
                   )
          ),
          
          fluidRow(
            column(width=4,
                   plotOutput("roc4")
                   ),
            column(width=4,  
                   plotOutput("roc5")
                   ),
            column(width=4,
                   plotOutput("roc6")
                   )
          )
        )
      
    )
     
  )
)
server <- function(input, output, session){
    output$plot1 <- renderPlot({
          ggplot(tloss, aes( x = epoch, y = train_loss))+ geom_line()+ ggtitle("Train Loss Function")
          })
    output$plot2 <- renderPlot({
        ggplot(vloss, aes( x = epoch, y = valid_loss))+ geom_line()+ ggtitle("Valid Loss Function")
    })
    
    output$plot4 <- renderPlot({
        xtbloss <- tbloss %>% filter(epoch == input$epoch)
        ggplot(xtbloss, aes( x = batch, y = loss))+ geom_line() + ggtitle("Valid Loss")
    })

    output$plot3 <- renderPlot({
        xvbloss <- vbloss %>% filter(epoch == input$epoch)
        ggplot(xvbloss, aes( x = batch, y = loss))+ geom_line() + ggtitle("Train Loss")
    })
    
    output$plot6 <- renderPlot({
        xtbloss <- tbloss %>% filter(epoch == input$epoch2)
        ggplot(xtbloss, aes( x = batch, y = acc))+ geom_line() + ggtitle("By Epoch")
    })

    output$plot5 <- renderPlot({
       ggplot(vloss, aes( x = epoch, y = valid_acc))+ geom_line()+ ggtitle("By Batch")
    })
    
    output$roc1 <- renderPlot({
      pROC_any <- roc(
            target$any, predicted$any,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_any <- ci.se(pROC_any)

      
       title(main="Any")
       plot(sens.ci_any, type="shape", col="lightblue")

    })
    
    output$roc2 <- renderPlot({
      pROC_epidural <- roc(
            target$epidural, predicted$epidural,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_epidural <- ci.se(pROC_epidural)

      
       title(main="Epidural")
       plot(sens.ci_epidural, type="shape", col="lightblue")

    })
    
    output$roc3 <- renderPlot({
      pROC_intraparenchymal <- roc(
            target$intraparenchymal, predicted$intraparenchymal,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_intraparenchymal <- ci.se(pROC_intraparenchymal)

      
       title(main="Intraparenchymal")
       plot(sens.ci_intraparenchymal, type="shape", col="lightblue")

    })

    output$roc4 <- renderPlot({
      pROC_intraventricular <- roc(
            target$intraventricular, predicted$intraventricular,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_intraventricular <- ci.se(pROC_intraventricular)

      
       title(main="Intraventricular")
       plot(sens.ci_intraventricular, type="shape", col="lightblue")

    })

    output$roc5 <- renderPlot({
      pROC_subarachnoid <- roc(
            target$subarachnoid, predicted$subarachnoid,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_subarachnoid <- ci.se(pROC_subarachnoid)

      
       title(main="Subarachnoid")
       plot(sens.ci_subarachnoid, type="shape", col="lightblue")

    })

    output$roc6 <- renderPlot({
      pROC_subdural <- roc(
            target$subdural, predicted$subdural,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_subdural <- ci.se(pROC_subdural)

      
       title(main="Subdural")
       plot(sens.ci_subdural, type="shape", col="lightblue")

    })
            
    }
shinyApp(ui = ui, server = server)


```

## Model-2 Metrics (Multilabel)

```{r}

ui <- fluidPage(
  tabsetPanel(
    tabPanel("Loss Function",
             splitLayout(
               plotOutput("plot1")
               ,
               plotOutput("plot2")          
               )
             
             ),
    tabPanel("Loss by Epoch",
             sidebarLayout(
              sidebarPanel(
                selectInput("epoch","No.Epoch", c(1,2,3,4,5,6,7,8,9,10,11,12),multiple = FALSE)
              ),
              mainPanel(
                splitLayout(
                  plotOutput("plot3"),
                  plotOutput("plot4")
                   )
                )
             )
             
        ),
     tabPanel("Accuracy",
         sidebarLayout(
           sidebarPanel(
             selectInput("epoch2","No.Epoch", c(1,2,3,4,5,6,7,8,9,10,11,12),multiple = FALSE)
           ),
           mainPanel(
             
              splitLayout(
                  plotOutput("plot5"),
                  plotOutput("plot6")
                   )
           )
         )
               
      ),
    tabPanel("Test Results",
        verticalLayout(
          fluidRow(
            
            column(width=4,
                   plotOutput("roc1")
                   ),
            column(width=4,  
                   plotOutput("roc2")
                   ),
            column(width=4,
                   plotOutput("roc3")
                   )
          ),
          
          fluidRow(
            column(width=4,
                   plotOutput("roc4")
                   ),
            column(width=4,  
                   plotOutput("roc5")
                   ),
            column(width=4,
                   plotOutput("roc6")
                   )
          )
        )
      
    )
     
  )
)
server <- function(input, output, session){
    output$plot1 <- renderPlot({
          ggplot(tloss2, aes( x = epoch, y = train_loss))+ geom_line()+ ggtitle("Train Loss Function")
          })
    output$plot2 <- renderPlot({
        ggplot(vloss2, aes( x = epoch, y = valid_loss))+ geom_line()+ ggtitle("Valid Loss Function")
    })
    
    output$plot4 <- renderPlot({
        xvbloss2 <- vbloss2 %>% filter(epoch == input$epoch)
        ggplot(xvbloss2, aes( x = batch, y = loss))+ geom_line() + ggtitle("Valid Loss")
    })

    output$plot3 <- renderPlot({
        xtbloss2 <- tbloss2 %>% filter(epoch == input$epoch)
        ggplot(xtbloss2, aes( x = batch, y = loss))+ geom_line() + ggtitle("Train Loss")
    })
    
    output$plot6 <- renderPlot({
        xvbloss2 <- vbloss2 %>% filter(epoch == input$epoch2)
        ggplot(xvbloss2, aes( x = batch, y = acc))+ geom_line() + ggtitle("By Epoch")
    })

    output$plot5 <- renderPlot({
       ggplot(vloss2, aes( x = epoch, y = valid_acc))+ geom_line()+ ggtitle("By Batch")
    })
    
    output$roc1 <- renderPlot({
      pROC_any <- roc(
            target2$any, predicted2$any,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_any <- ci.se(pROC_any)

      
       title(main="Any")
       plot(sens.ci_any, type="shape", col="lightblue")

    })
    
    output$roc2 <- renderPlot({
      pROC_epidural <- roc(
            target2$epidural, predicted2$epidural,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_epidural <- ci.se(pROC_epidural)

      
       title(main="Epidural")
       plot(sens.ci_epidural, type="shape", col="lightblue")

    })
    
    output$roc3 <- renderPlot({
      pROC_intraparenchymal <- roc(
            target2$intraparenchymal, predicted2$intraparenchymal,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_intraparenchymal <- ci.se(pROC_intraparenchymal)

      
       title(main="Intraparenchymal")
       plot(sens.ci_intraparenchymal, type="shape", col="lightblue")

    })

    output$roc4 <- renderPlot({
      pROC_intraventricular <- roc(
            target2$intraventricular, predicted2$intraventricular,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_intraventricular <- ci.se(pROC_intraventricular)

      
       title(main="Intraventricular")
       plot(sens.ci_intraventricular, type="shape", col="lightblue")

    })

    output$roc5 <- renderPlot({
      pROC_subarachnoid <- roc(
            target2$subarachnoid, predicted2$subarachnoid,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_subarachnoid <- ci.se(pROC_subarachnoid)

      
       title(main="Subarachnoid")
       plot(sens.ci_subarachnoid, type="shape", col="lightblue")

    })

    output$roc6 <- renderPlot({
      pROC_subdural <- roc(
            target2$subdural, predicted2$subdural,
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


      sens.ci_subdural <- ci.se(pROC_subdural)

      
       title(main="Subdural")
       plot(sens.ci_subdural, type="shape", col="lightblue")

    })
            
    }
shinyApp(ui = ui, server = server)

```

## Results and Conclusions
- Batches of 15 or 10 images gave better results
- The prototype shows that two models are a good approach
- The use of the function loss in the validation phase is not a good option
- The accuracy and other metrics are necessary to evaluate the model
- The number of images use in "Any" category shows a good metric for sampling the other clases
- It is important to include feature enhancing on the model
- The prototype is ready to go into second phase 
    - Modification Data Augmentation and Sampling
    - Fine tunning 
    - Including InceptionV3

## Leasons Learned 

- Good resarch can help avoid mistakes and learn new techniques to include in our models
- Good presentation of metrics helps in the decision process
- Pytorch is very helpful to include controls in the model
- Learn about memory usage of GPUs is important to design Network architectures


